---
title: "Lab1"
author: "Kelsie Fronheiser"
date: "1/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(here)
library(sf)
library(tmap)
library(tidyr)
```


#### Read in SF trees data
```{r}
sf_trees <- read_csv(here("data", "sf_trees", "sf_trees.csv"))
```

#### Basic wrangling reminders

Find top five highest observations of trees by legal status & do some wrangling and make a graph
```{r}
top_5_status <- sf_trees %>% 
  count(legal_status) %>% 
  #group by, summarize, and n in one function
  drop_na(legal_status) %>% 
  rename(tree_count = n) %>% 
  #new name = orignal name
  relocate(tree_count) %>% 
#move location of column, this moved it to the front
  slice_max(tree_count, n = 5)
# picked out top 5 highest values of tree_count

```

Make a graph of top 5 observations by legal status
```{r}
ggplot(data = top_5_status, aes(x = fct_reorder(legal_status, tree_count), y = tree_count)) +
  geom_col() +
  labs(x = "Legal Status", y = "Tree Count") +
  coord_flip() +
  theme_minimal()

#fct_reorder is recording thelegal status  column locations on the graph by the tree count

```
### More data wrangling

Only keep obs for blackwood acacia tree

```{r}
blackwood_acacia <- sf_trees %>% 
  filter(str_detect(species, "Blackwood Acacia")) %>% 
#keep obs where certain string is detected anywhwere in that obs
  select(legal_status, date, latitude, longitude)
#selecting which columns show up in dataset

ggplot(data = blackwood_acacia, aes(x = longitude, y = latitude)) +
  geom_point()
```

## Using tidyr :: separate() and unite() functions that are useful for combining or seperating columns

```{r}
sf_trees_sep <- sf_trees %>% 
  separate(species, into = c("spp_scientific", "spp_common"), sep = "::")
```

Example: tidyr :: unite()

```{r}
sf_tress_unite <- sf_trees %>% 
  unite("id_status", tree_id : legal_status, sep = "_cool!_")
```




